---
title: "wrangling2.Rmd"
author: "mikec964"
date: "August 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
```


# Overview
This is an exercise for the Springboard Data Foundations course. In each step
we'll look at the data from the Titanic and fill in missing data.

0. Load the data
#. Port of embarcation
#. Age
#. Lifeboat
#. Cabin


# Load the data
We're going to load the original Excel spreadsheet, *titanic3.xls*. We'll
save an as-is copy as *titanic_original.csv*.
```{r}
titanic_orig <- read.csv("titanic3_original.csv")
titanic <- titanic_orig
```


# Port of embarcation
The embarked column has some missing values, which are known to correspond to
passengers who actually embarked at Southampton. Find the missing values and 
replace them with S.

FWIW, some rows to fix: 169, 284
```{r}
print(titanic$embarked[168:170])
for(p in 1:length(titanic$embarked)) {
  if(titanic$embarked[p] == "") { titanic$embarked[p] <- "S" }
}
print(titanic$embarked[168:170])
```


# Age
Calculate the mean of the Age column and use that value to populate the missing 
values.

FWIW, some rows to fix: 16, 41
```{r}
print(titanic$age[15:17])
age.m <- mean(titanic$age, na.rm=TRUE)
for(p in 1:length(titanic$age)) {
  if(is.na(titanic$age[p])) { titanic$age[p] <- age.m }
}
print(titanic$age[15:17])
```


